Procedural generation of levels fot he angry birds videogame using evolutionary computation\\

### 2019-05-14 IEEE_ConferenceOnGames/article_english.tex
Update to article 15/5/2019 00:44

\section{Introduction}
Procedural content generation is an interesting subject on the entertainment
industry, primarily in the video game development because of the easines and
quickness provided to generate content for diferent video game projects [ref] .
In this paper an evolutionary content generation system is proposed to generate
levels for the popular Angry Birds video game
\cite{RovioEntertainmentCorporation2009}, taking inspiration in concepts of
open-ended evolution.

For this project a procedural generation content system is proposed for the game
angry birds, angry birds is a game developed by Rovio Entrtainment in which the
objective is to destroy structures placed on the map as well as eliminating
green pig-like entities in order to obtain the highest score possible, the game
uses a gravity and collition system that allows the structures to fall and be
destroyed by means of damage as well as the use of three diferent construction
materials \textit{ice}, \textit{wood} and \textit{stone} each one with different
resistances ranging from fragile, normal and sturdy in that order, the game
mechanic is to use a slingshot to shoot a certain number of birds with different
abilities to destroy the structures and the objectives, it is possible to obtain
a high score by doing the most destruction as possible with the least birds as
you can. \cite{RovioEntertainmentCorporation2009}

@@

% What is the problem?
% How we solve it? (summary)   
% What is learned?

In Section 2, a background of the subjects used on this project will be
presented, this subjects, while not all being used on by different authors on
their works provide a great help on the development of the proposed system,
% Jaime, remember there is no Background need it in this type of paper 
% Maybe that will be the introduction - MG

Section 3, contains a quick overview of some aspects used by different authors
in order to tackle this problem, their design and basic functionality, in
Section 4 the description of the current state of the project is explained, from
the basis of the development to the diferent menthods that have been used to
progress on the different sections of the project, 
% Normaly there is no "Future Work" section is put together with Conclusions - MG
in Section 5 we explain the
future work of the project, where we are in the development phase and how the
future implementations are to further improve the project and finally, in
Section 6 we present our conclusions

\section{Background}

In this Section we provide a quick overview of the literature about the areas
that provide the basis to the development of the current project, this section
is divided in three parts, the fisrt one gives us an explanation of what
procedural content generation entails, the second one provides an explanation on
genetic algorithms and how can they be used to solve optimization problems, and
finally, the third one provides us with an overview of Open-Ended Evolution and
% use just one form, I prefer with no first caps open-ended evolution and let's 
% use an abbreviation OEE - MG

what it means to have an Open-Ended Evolutionary algorithm.
% For this conference I think we do not need to define PCG, let's think 
% every one knows what it is. - MG

\subsection{Procedural content generation}

Procedural content generation (PCG) denotes the way to automatically generate
content by algorithms instead of generating the same content by manual means
such as with different personnel.

@@

\subsection{Genetic algorithm}
Is a method developed by John Holland on 1975 \cite{Holland1975} the genetic
algorithm (GA) is an optimization and search technique based on the principles
of genetic and natural selection, it allows a population composed of many
individuals, which by themselves represent a possible solution to the problem,
to evolve under specific selection rules in order to obtain the maximum fitness.

In other words a genetic algorithm is a method to solve optimization problems by
creating a population that will represent a possible solution and repeatedly
doing genetic operations as selection, crossover and mutation on each member of
the population in order to generate the members of the population for the next
generation that would be closer to the best solution of the problem than their
parents, the cycles in this method where the operations are made are called
generations and by moving over this the population “evolves” o in other words,
they get closer to the best possible solution  to a specific problem.

@@

@@

\subsection{Open-Ended Evolution}\label{AA}
Open-ended evolution (OEE) is a variant of genetic evolution algorithm in which
the objective differs from the regular genetic algorithm one, in this case the
final objective is not to reach a final state of the evolution but to continue
evolving and creating new types of entities more complex on each new generation,
these same entities have the possibility to evolve not only inside the frame of
their type but create entirely new groups. \cite{Standish2003}

@@

@@

The Angry Birds level generation track is a competition where the objective is 
to generate levels that are \textit{fun} and \textit{enjoyable} however they 
also have to be challenging but imposible however the rules for the competition 
specify that a configuration file must be used to
generate the levels of the game in a limited amount of time, the infornmation on
the configuration file is first the number of \textbf{\textit{diferent}} levels
that must be generated, the second value is the diferent materials \textit{not
allowed} to be used, for instance a group of generated levels must not have
circles which composition material is ice or stone moreover it des not have to
integrate a triangle piece with ice, stone or wood material so this combinations
must be removed from the pool of options, the third configuration value is the
\textit{number of pigs} that must be placed in a level, and finally the
\textit{limit} of time that the system has to create the required amount of
levels.

@@

The system as been able to generate levels that fulfill the requirements
 provided by the parameters file, however a way to evaluate that the levels
 generated are indeed \textit{"fun"} or \textit{"interesting"} is still needed,
 at this point we have two was to evaluate the results:
 \begin{itemize}
    \item User survey, in case a survey is used as an evaluation tool then first
    the system has to generate the levels and a group of people must participate
    by playing the levels and answering a set of questions afterwards, the
    questions themselves must try to capture the feeling of a person towards the
    game like questioning if a level was too easy or too dificult to solve, if
    the generated structures look interesting or if the level required them to
    strategize how to use the provided birds to solve the level, the survey must
    be applied to a equal quantity of people of three diferent game skill
    levels,\textit{"non video game players"}, \textit{"casual video game
    players"} and \textit{"avid video game players"}, this way we can evaluate a
    level or the entire system by knowing if all skill levels did not enjoyed a
    level because it was too simple or easy.
    \item By using a naive agent, if a naive agent is used to play the levels we
    can obtain data of how the agent solves the levels placed before them, and
    the way to evaluate if the level is or not \textit{"interesting"} will be by
    checking how easily the aggent is able to solve a level, the point here is
    to have a level that cannot be solved by a single movement but to have one
    that a human player has to think a while to solve tryingto place the
    difficulty of a level just at the middle point between easy and hard
    difficulty levels.
\end{itemize}

>>>>

\section{Introduction}
Procedural content generation is an interesting subject on the entertainment
industry, primarily in the video game development because of the easines and
quickness provided to generate content for different video game projects [ref] .
In this paper an evolutionary content generation system is proposed to generate
levels for the popular Angry Birds video game
\cite{RovioEntertainmentCorporation2009}, taking inspiration in concepts of
open-ended evolution.

For this project a procedural generation content system is proposed for the game
angry birds, angry birds is a game developed by Rovio Entrtainment in which the
objective is to destroy structures placed on the map as well as eliminating
green pig-like entities in order to obtain the highest score possible, the game
uses a gravity and collition system that allows the structures to fall and be
destroyed by means of damage as well as the use of three different construction
materials \textit{ice}, \textit{wood} and \textit{stone} each one with different
resistances ranging from fragile, normal and sturdy in that order, the game
mechanic is to use a slingshot to shoot a certain number of birds with different
abilities to destroy the structures and the objectives, it is possible to obtain
a high score by doing the most destruction as possible with the least birds as
you can. \cite{RovioEntertainmentCorporation2009}

@@

% What is the problem?
Procedural content generation provides a peculir case of study because of the way
content has to be made for a each particular type of game, in this particular case 
the restrictions that the game and the competition track have provide an interesting 
challenge for a evolutionary algorithm because we have to tanke into account the 
different pieces that can be used, their interactions with one another because of 
gravity and the possible effects the trajectory of a shot from a player can affect 
all of this.

% How we solve it? (summary)   
In order to tackle this case of study we proposed to use a genetic algorithm capable 
of creating content, more specifically levels for the angry birds video game in which
the generated levels satisfy the objectives of being \textit{interesting} and 
\textit{playable}.

% What is learned?
After concluding this paper we expect to identify the places were a evolutionary 
algorithm can be used alongside procedural content generation applied to video games
in order to be able to clarify in which areas a evolutionary algorithm can used 
to further improve video game development cycles.

As mentioned above in this paper we use procedural content generation (PCG) 
which denotes the way to automatically generate
content by algorithms instead of generating the same content by manual means
such as with different personnel.

@@

Furthermore we use genetic algorithms which is a method developed by 
John Holland on 1975 \cite{Holland1975} the genetic
algorithm (GA) is an optimization and search technique based on the principles
of genetic and natural selection, it allows a population composed of many
individuals, which by themselves represent a possible solution to the problem,
to evolve under specific selection rules in order to obtain the maximum fitness.

In other words a genetic algorithm is a method to solve optimization problems by
creating a population that will represent a possible solution and repeatedly
doing genetic operations as selection, crossover and mutation on each member of
the population in order to generate children for the next
generation that would be closer to the best solution of the problem than their
parents, the cycles in this method where the operations are made are called
generations and by moving over this the population “evolves” o in other words,
they get closer to the best possible solution  to a specific problem.

@@

And finally we use our own interpretation of an Open-ended evolution (OEE) 
algorithm which is a variant of genetic evolution algorithm where
the objective differs from the regular genetic algorithm one, in this case the
final objective is not to reach a final state of the evolution but to continue
evolving and creating new types of entities more complex on each new generation,
these same entities have the possibility to evolve not only inside the frame of
their type but create entirely new groups. \cite{Standish2003}

%In Section 2, a background of the subjects used on this project will be
%presented, this subjects, while not all being used on by different authors on
%their works provide a great help on the development of the proposed system,
% Jaime, remember there is no Background need it in this type of paper 
% Maybe that will be the introduction - MG

Section 2, contains a quick overview of some aspects used by different authors
in order to tackle this problem, their design and basic functionality, in
Section 3 the description of the current state of the project is explained, from
the basis of the development, modification to the original proposed method as 
well as different approaches that were taken to solve the problem.
% to the different methods that have been used to
%progress on the different sections of the project, 
% Normaly there is no "Future Work" section is put together with Conclusions - MG
finally in Section 4 we provide our conclusions about the development of this
paper as well as improvements that can help to refine the proposed work.

%we explain the future work of the project, where we are in the development phase and how the
%future implementations are to further improve the project and finally, in
%Section 6 we present our conclusions

%\section{Background}
%
%In this Section we provide a quick overview of the literature about the areas
%that provide the basis to the development of the current project, this section
%is divided in three parts, the fisrt one gives us an explanation of what
%procedural content generation entails, the second one provides an explanation on
%genetic algorithms and how can they be used to solve optimization problems, and
%finally, the third one provides us with an overview of open-ended evolution and
% use just one form, I prefer with no first caps open-ended evolution and let's 
% use an abbreviation OEE - MG

%what it means to have an open-ended evolution algorithm.
% For this conference I think we do not need to define PCG, let's think 
% every one knows what it is. - MG

%\subsection{Procedural content generation}

%Procedural content generation (PCG)

%\subsection{Genetic algorithm}

@@

%\subsection{Open-Ended Evolution}\label{AA}

@@

%                                                                           /|\
% Combined a few other things to give a litte more detail to this sections   |
%                                                                           \|/

@@

The Angry Birds level generation track is a competition where the objective is 
to generate levels that are \textit{fun} and \textit{enjoyable} however they 
also have to be challenging but imposible however the rules for the competition 
specify that a configuration file must be used to
generate the levels of the game in a limited amount of time, the infornmation on
the configuration file is first the number of \textbf{\textit{different}} levels
that must be generated, the second value is the different materials \textit{not
allowed} to be used, for instance a group of generated levels must not have
circles which composition material is ice or stone moreover it des not have to
integrate a triangle piece with ice, stone or wood material so this combinations
must be removed from the pool of options, the third configuration value is the
\textit{number of pigs} that must be placed in a level, and finally the
\textit{limit} of time that the system has to create the required amount of
levels.

@@

The system as been able to generate levels that fulfill the requirements
 provided by the parameters file, however a way to evaluate that the levels
 generated are indeed \textit{"fun"} or \textit{"interesting"} is still needed,
 at this point we have two was to evaluate the results:
 \begin{itemize}
    \item User survey, in case a survey is used as an evaluation tool then first
    the system has to generate the levels and a group of people must participate
    by playing the levels and answering a set of questions afterwards, the
    questions themselves must try to capture the feeling of a person towards the
    game like questioning if a level was too easy or too dificult to solve, if
    the generated structures look interesting or if the level required them to
    strategize how to use the provided birds to solve the level, the survey must
    be applied to a equal quantity of people of three different game skill
    levels,\textit{"non video game players"}, \textit{"casual video game
    players"} and \textit{"avid video game players"}, this way we can evaluate a
    level or the entire system by knowing if all skill levels did not enjoyed a
    level because it was too simple or easy.
    \item By using a naive agent, if a naive agent is used to play the levels we
    can obtain data of how the agent solves the levels placed before them, and
    the way to evaluate if the level is or not \textit{"interesting"} will be by
    checking how easily the aggent is able to solve a level, the point here is
    to have a level that cannot be solved by a single movement but to have one
    that a human player has to think a while to solve tryingto place the
    difficulty of a level just at the middle point between easy and hard
    difficulty levels.
\end{itemize}
